---
title: "153project"
output:
  pdf_document: default
  html_document: default
---

```{r}
library(astsa)
library(TSA)
library(tsoutliers)
#library(randtests)
```

2.1 Set up and EDA
```{r}
# setwd("/Users/furonghuang/Documents/Study materials/Statistics/Time Series/Project")
hotel.raw = read.csv("monthly-hotel-occupied-room-av-6.csv")
colnames(hotel.raw)=c("Month", "Rooms")
hotel = hotel.raw[-169,]
hotel.ts = ts(hotel[,2], start=c(1963, 1), end = c(1976,12), frequency = 12)
hotel.train = ts(hotel[,2], start=c(1963, 1), end = c(1974,12), frequency = 12) # select 1963-1974 as train set
hotel.train.df = hotel[1:144, ]
hotel.test = ts(hotel[145:168, 2], start=c(1975, 1), end = c(1976,12), frequency = 12) # leave last 2 yrs as test set
hotel.test.df = hotel[145:168, ]
hotel$Month = as.Date(paste(as.character(hotel$Month), "-01", sep=""),
                      format = "%Y-%m-%d")
# in order to show yearly tick on the axis, have to use data.frame for plotting
plot(Rooms~Month, data= hotel, type="l", 
     main="Time Series: Monthly data of hotel occupied room from 1963-01 to 1976-12")
axis.Date(side=1, at=seq.Date(min(hotel$Month),max(hotel$Month), by="1 year"))
```

Test for outliner
```{r}
tso(hotel.train, types = c("TC", "AO", "LS", "IO", "SLS"))
```


Chasing stationarity
```{r}
l.hotel.train = log(hotel.train)
dl.hotel.train = diff(l.hotel.train)
ddl.hotel.train = diff(dl.hotel.train,12)
plot.ts(cbind(l.hotel.train, dl.hotel.train, ddl.hotel.train))
```

2.2 Spectral analysis 

detrend by linear regression, residual plot
```{r}
t = 1:length(l.hotel.train)
fit = lm(l.hotel.train ~ t)
detrend.hotel.train = fit$residuals 
# to use spec.pgram() for periodogram, we need stationary, so use the detrend data. However, for spec.pgram(), it automatically detrend the data for you, so fitting the log data is also fine
plot.ts(detrend.hotel.train, main="Detrend hotel")
```

Periodogram without smoothing and tapering
```{r}
spec.pgram(detrend.hotel.train, taper = 0)
```

Use daniell kernel to smooth the periodogram
```{r}
spec.pgram(detrend.hotel.train, kernel("daniell", 1), taper=0)
```

To avoid the flat region at the peaks, use modified.daniell kernel
```{r}
spec.pgram(detrend.hotel.train, kernel("modified.daniell", 1), taper=0)
```

To reduce the side lobes, use tapering
```{r}
spec.pgram(detrend.hotel.train, kernel("modified.daniell", 1), taper=0.2)
```

Find the key frequencies
```{r}
pgram = spec.pgram(detrend.hotel.train, kernel("modified.daniell", c(1, 1, 1)), taper=0.2)
key_freq_ind = c(1, which(diff(sign(diff(pgram$spec)))==-2) + 1)
key_freq = pgram$freq[key_freq_ind]
abline(v=key_freq, lty=2)
# notice that the periodogram should range from [0, 1/2], sometimes if you use the log data(which retain the TS format), spec.pgram() will show key freq as 1,2,3... for convenience. if it is monthly data, we need to divide those key freq by 12 to get back the range [0, 1/2]
```

Parametric way to find the periodogram and comparision to Nonparametric way.
```{r}
spec.pgram(detrend.hotel.train, kernel("modified.daniell", c(1, 1, 1)), taper=0.2)
pgram_ar = spec.ar(detrend.hotel.train, plot=F) # plot the parametric spectral estimation as red, it has the same peaks as nonparametric method, good
lines(pgram_ar$freq, pgram_ar$spec, lty=2, col="red")
```
Parametric and Nonparametric yeild the same key frequencies

Check the model 
```{r}
top_freq = key_freq[order(pgram$spec[key_freq_ind], decreasing = T)][1:5]
periodic_terms = do.call(cbind, lapply(top_freq, function(freq) {
  cbind(cos(2 * pi * freq * t), sin(2 * pi * freq * t))
})) # no need to add columns of 1 and t because the fn lm() will accomodate the trend
df = data.frame(l.hotel.train, t, periodic_terms) # change: fit the original data instead of the log data, linear regression will accomodate the trend.
fit_final = lm(l.hotel.train ~ ., df) # first 2 cols are for linear trend, next cols are cos&sin terms, this linear regression will find coeffs for linear trend and periodic trend(thoes Ui, Uj)

plot(t, l.hotel.train, type="l", main ="Original(black) and Fitted(red)") # plot the original time series
lines(t, fit_final$fitted.values, lty=2, col="red")
```

residual of the actual data and the model data
```{r}
plot(t, fit_final$residuals, type = "l", main="Residues of model")
```
The residuals are nearly equally spread on two sides of y=0 line and have a constant mean. 

Predict the test set
```{r}
t_new = (tail(t, 1) + 1):(tail(t, 1) + 24)
periodic_terms_new = do.call(cbind, lapply(top_freq, function(freq) {
  cbind(cos(2 * pi * freq * t_new), sin(2 * pi * freq * t_new)) # key freqs are the same for the whole data, so use them to generate new periodic terms
}))
df_new = data.frame(t_new, periodic_terms_new)
colnames(df_new) = colnames(df)[-1]
hotel.pred.periogram = predict.lm(fit_final, newdata=df_new, interval="prediction", level=.95) # calc prediction based on the model coeffs and the new data points, note this is the log data, need to convert it back

plot(c(t, t_new), hotel$Rooms, type="l", xlim=c(0, tail(t_new, 1)), ylim=c(500, 1200), xlab = "t (month)", ylab = "hotel occupancy", main="Prediction of test set (1975-1976)") # black curve is actural values
lines(t, exp(fit_final$fitted.values), lty=2, col="red")
lines(t_new, exp(hotel.pred.periogram[, "fit"]), col="blue") # blue is prediction for test set 

```

Real test value compared to predicted value
```{r}
plot(t_new, hotel.test, type = "l", ylim = c(700,1200))
lines(t_new, exp(hotel.pred.periogram[, "fit"]), col="blue") # blue is prediction for test set 
```

add CI
```{r}
plot(t_new, hotel.test, type = "l", ylim = c(700,1200), main="Zoom in the prediction of test set", xlab = "t (month)", ylab = "hotel occupancy")
lines(t_new, exp(hotel.pred.periogram[, "fit"]), col="blue") # blue is prediction for test set 
matlines(t_new, exp(hotel.pred.periogram[, 2:3]), col = "purple", lty=3) #purple is CI
```

MSE of the test set
```{r}
mean((exp(hotel.pred.periogram[, "fit"])-hotel.test)^2)
```

Predict the future 2 years (1977-1978)
```{r}
t_pred = (tail(t_new, 1) + 1):(tail(t_new, 1) + 24)
plot(c(t, t_new), hotel$Rooms, type = "l", xlim = c(0, tail(t_pred,1)), ylim = c(500,1200), xlab = "t (month)", ylab = "hotel occupancy", main="Prediction of future 2 years (1977-1978)")
lines(t_pred, exp(hotel.pred.periogram[,"fit"]), col="blue")
matlines(t_pred, exp(hotel.pred.periogram[,2:3]), col='purple', lty = 3)
```


2.3 Fit SARIMA
ACF 
```{r}
acf2(dl.hotel.train,50) # see a seasonal pattern every 12 months
acf2(ddl.hotel.train, 50)
```
From the ACF of the first differencing data, see a yearly seasonal pattern

Estimate parameter
```{r}
eacf(dl.hotel.train)
```
eacf of the first differencing data gives no information, check the eacf of the seasonal differencing data to get further guess of the model parameter(seasonal)



Fit model
Tests several possible parameter
```{r}
sarima(l.hotel.train, 0,1,3, 0,1,1, 12) # AIC -6.92
```

```{r}
sarima(l.hotel.train, 0,1,4, 0,1,1, 12) # AIC -6.9
```

```{r}
sarima(l.hotel.train, 1,1,1, 0,1,1, 12) #AIC -6.8
```

```{r}
sarima(l.hotel.train, 0,1,2, 0,1,1, 12) # AIC -6.8
```

```{r}
sarima(l.hotel.train, 0,1,1, 1,1,1, 12)# AIC -6.75
```

```{r}
sarima(l.hotel.train, 0,1,6, 2,1,1, 12) # AIC -7.00
```

```{r}
sarima(l.hotel.train, 0,1,6, 0,1,3, 12) #AIC -7.00
```

Get the best model
```{r}
sarima(l.hotel.train, 0,1,6, 0,1,1, 12)
# by several trials of parameters, will write more choices of parameters for comparison purpose
```

Verify the model: 

Take the last two years as the test set
```{r}
#hotel.train = ts(hotel.ts[1:144], start=c(1963, 1), end = c(1974,12), frequency = 12 ) # the train set
#l.hotel.train = log(hotel.train)
#hotel.test = hotel[145:168,] # test set
```

predict the "unseen" test data for model verify purpose
```{r}
test.arima = sarima.for(l.hotel.train, 24, 0,1,6, 0,1,1, 12)
```

Plot the predicted value of the test data along with the test data
```{r}
exp.test.arima = exp(test.arima$pred)

plot(hotel.train, xlim=c(1963, 1977), ylim=c(400,1300), 
     main="Time Series predict vs. observed: Monthly data of hotel occupied room ")
lines(exp.test.arima, col="red") # red line denotes the predicted value
lines(hotel.test, lty = 2)

```

Comparison of test set and observed set
```{r}
plot(hotel.test)
lines(exp.test.arima, col="red") # red line denotes the predicted value

```

Mean square error
```{r}
mean((exp.test.arima - hotel.test)^2) # 274.4558
```



Prediction
```{r}
pred.hotel = sarima.for(l.hotel.train, 24,  0,1,6, 0,1,1, 12)
```

Plot back to the original scale
```{r}
plot(hotel.ts, xlim=c(1963, 1979), ylim=c(400,1300), 
     main="Time Series predict vs. observed: Monthly data of hotel occupied room ")
lines(exp(pred.hotel$pred), col ="red")
# add CI for predicted value
lines(exp(pred.hotel$pred + pred.hotel$se), col="blue", lty =3) # upper CI
lines(exp(pred.hotel$pred - pred.hotel$se), col="green", lty = 3) # lower CI
```



